---
title: "2_13"
author: "Alex Tomack"
date: "2023-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)

nba<- read_rds("https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/3_Data_Wrangling/data/nba_players_2018.Rds?raw=true")
gms<-read_rds("https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/4_Uni_Multivariate/data/game_summary.Rds?raw=true")
```

What is confidence?

- In frequent stats:
  - How often your concliusion would be correct if you were able to run the experiment many times
  - How often your conclusion would be correct if you were able to observe the world many times
- Research question: Are NBA players from Tennessee better at shooting free throws than players from UVA?

```{r}
glimpse(nba%>%select(org, pctFT)) # look at the types of variables we're working with

summary(nba %>%select(pctFT, org)) # check how many NAs we have. 
```
Now that we've looked, let's do some univariate visualization
```{r}
# univariate of pctFT
nba %>% 
  ggplot(aes(x=pctFT))+
  geom_density()

# univariate of org
nba %>%
  count(org) %>%
  ggplot(aes(x=n, y= reorder(org, n)))+
  geom_bar(stat="identity")
```

Multivariate visualization 
```{r}
nba %>%
  filter(org %in% c("Tennessee", "Virginia")) %>%#same as filter(org=="Tennessee" | org=="Virginia")
  group_by(org)%>%
  summarise(meanFT=mean(pctFT)) %>%
  ggplot(aes(x=org, y=meanFT))+
  geom_bar(stat="identity")

#Good visualization
nba %>%
  filter(org %in% c("Tennessee", "Virginia")) %>%#same as filter(org=="Tennessee" | org=="Virginia")
  ggplot(aes(x=org, y=pctFT))+
  geom_boxplot()
```
Uncertainty

- Are players from Tennessee better at free throws than players from UVA?
- Populations vs samples
- Big philosophical step back
  - we live in a stochastic universe
- What does better mean?
  - Theory: An innate quality in greater abundance
  - Prediction: If we had to bet on who scores more FTs, who do we choose?
- If the universe is inherently stochastic, we are inherently uncertain. 
  - How do we measure this?
    - Run 100 experimental seasons 
    - Record FT percentage for players from UVA and UT for each season
    - Calculate how many times UT players have better percentage than UVA players
  - 90 seasons out of 100-- 90% confident/certain
  - can never be 100% confident.
  - But, running 100 experimental seasons is impossible.
    - What we can do, is take 1 season of basketball but sample it randomly.
      - Bootstrap sampling
    - Theory: By mimicking the sampling process, we can simulate a God experiemtn
    - Practice: sample_n(size, replace)
    
Bootstrap Demo Step 1

- One randomly sampled player via sample_n(size, replace)
  - Size: how many samples (from 1 to all observations)
  - replace: whether to put the sample back (true/false)
  
# Bootstrap Example
```{r}
set.seed(123)
nba %>%
  select(org, pctFT) %>%
  sample_n(size=1, replace=T)

# Get a real simulated season
simSeason1<-nba %>%
  select(org, pctFT) %>%
  sample_n(size=nrow(nba), replace=T) # take 530 observations and put back into data set

simSeason1 %>%
  filter(org %in% c("Tennessee", "Virginia")) %>%
  group_by(org) %>%
  summarise(meanFT=mean(pctFT))
```

#repeating a chunk with "for()" loop
```{r}
bootstrap_result<- NULL #instantiating an object in memory
for(i in 1:100) {
  simSeason<-nba %>%
  select(org, pctFT) %>%
  sample_n(size=nrow(nba), replace=T) %>%
    mutate(bootstrap_number=i)
  
  bootstrap_result <- bootstrap_result %>%
    bind_rows(simSeason)
}

prepared_bootstrap<-bootstrap_result %>%
  filter(org %in% c("Tennessee", "Virginia")) %>%
  group_by(bootstrap_number, org) %>%
  summarise(meanFT=mean(pctFT), .groups="drop") %>%
  spread(key=org, value=meanFT) %>%
  mutate(diff=Tennessee-Virginia)

# calculate confidence
prepared_bootstrap %>%
  summarise(confidence=mean(diff > 0, na.rm=T)) # Find the mean of the diff values that are greater than 0
```
#Visualizing Uncertainty

- Method 1: plot the outcomes
```{r}
prepared_bootstrap %>%
  ggplot()+
  geom_density(aes(x=Tennessee), fill="orange", alpha=0.4)+
  geom_density(aes(x=Virginia), fill="brown", alpha=0.4) +
  geom_vline(xintercept=mean(prepared_bootstrap$Tennessee, na.rm=T),
             linetype="dashed", color="orange") +
  geom_vline(xintercept=mean(prepared_bootstrap$Virginia, na.rm=T),
             linetype="dashed", color="brown") +
  annotate(geom="text", x=mean(prepared_bootstrap$Tennessee, na.rm=T), y=Inf, label=round(mean(prepared_bootstrap$Tennessee, na.rm=T),2), angle=90, hjust=1, vjust=0)+
  annotate(geom="text", x=mean(prepared_bootstrap$Virginia, na.rm=T), y=Inf, label=round(mean(prepared_bootstrap$Virginia, na.rm=T),2), angle=90, hjust=1, vjust=0)
```
# Method 2: Plotting the estimates
```{r}
prepared_bootstrap %>%
  ggplot(aes(x=diff)) +
  geom_density() +
  geom_vline(xintercept = 0, linetype="dashed")
```
Where to Calculate the estimate

- First, we created a new dataset out of 100 simulated seasons
- Then, we calculate average FT % for TN and UVA for each sim
- Finally, we calculate proportion of times average is higher for TN

Interpreting confidence

- The closer your are to 50%, the less confident you are.


///////////////////////
# Review
//////////////////////
```{r}
require(tidyverse)
games<-read_rds("https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/4_Uni_Multivariate/data/game_summary.Rds?raw=true")
```

```{r}
games %>%
  group_by(isWin, yearSeason) %>%
  summarise(avgTO=mean(tov), 
            avgPts=mean(pts))%>%
 # spread(key=isWin, value=avgTO) or following
  pivot_wider(names_from = isWin, values_from = c("avgTO", "avgPts"))
```

To Bootstrap!
```{r}
bootstrap_result<-NULL
for(bsNumber in 1:100){
bootstrap_result<-games %>%
  sample_n(size=nrow(.), replace=T) %>% # the dot is the same thing as writing nrow(games). Dot just says to fill in the object which it's being linked to
  group_by(isWin)%>%
  summarise(avgTO=mean(tov)) %>%
  ungroup() %>% #best practice to ungroup after group
  bind_rows(bootstrap_result) # take the tibble and add it to the bottom of our Bootstrap result object. Adding two rows to the bottom 100 times.
  print(".")

}
bootstrap_result %>%
  spread(key=isWin, value=avgTO) %>%
  mutate(TO_diff=`TRUE`-`FALSE`) %>%
  summarise(conf=mean(TO_diff<0))
```




For final part of PS4-- how to look at a statistic for EVERY TEAM. Note tha addition of group_by(nameTeam) and group_by(..., nameTeam)

```{r}
set.seed(123)
bootstrap_result<-NULL
for(bsNumber in 1:100){
bootstrap_result<-games %>%
  group_by(nameTeam)%>%
  sample_n(size=n(), replace=T) %>% # the dot is the same thing as writing nrow(games). Dot just says to fill in the object which it's being linked to
  group_by(isWin, nameTeam)%>%
  summarise(avgTO=mean(tov),.groups="drop") %>%
  ungroup() %>% #best practice to ungroup after group
  bind_rows(bootstrap_result)
}

toplot<-bootstrap_result %>%
  pivot_wider(names_from=isWin,values_from=avgTO) %>%
  mutate(diff_TO=`TRUE`-`FALSE`)%>%
  group_by(nameTeam)%>%
  summarise(conf_TO=mean(diff_TO <0),
            avgDiff=mean(diffTO))

toplot %>%
  mutate(conf_binary=ifelse(conf_TO > 0.95, | conf_TO<0.05, 'Confident', 'Not Confident'))%>%
  ggplot(aes(x=avgDiff, y=reorder(nameTeam, avgDiff), fill=conf_binary))+
  geom_bar(stat="identity")
```