---
title: "2_20 Lecture"
author: "Alex Tomack"
date: "2023-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Regression and Conditional Analysis**

- Recall our discussion of conditional analysis
  - Conditional -> depends on 
  - Analyze with conditional means
  
  
**Reminder of the Process**

1. Determine the variable type
  - chr, fct for cat, dbl for cont
2. Type informs univariate analysis
3. Combinations of types informs conditional analysis


**What is Regression?**

- Conditional means for continuous data
  - But conditional means make a lot of mistakes. Can we do better? 
    - Yes! We use regression lines
    
**Regression**

- Calculating  a line that minimizes mistakes for every observation 
  - Y= a +bX
- All about reducing the error between each point and our regression line.
- The line is substantively meaningful. 


**Let's actually do some regression**
What is the relationship between SAT scores and median future earnings?

- The relationship might be positive
```{r}
require(tidyverse)
debt<-read_rds("https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/2_Intro_to_R/data/sc_debt.Rds?raw=true")

```
Step 1.) Look

- Look for missingness
  - Substantive: external validity 
  - Technical: cross validation wont work
  
```{r}
summary(debt %>% select(sat_avg, md_earn_wne_p6)) # lots of NAs, we'll have to drop them
```
Step 2.) Univariate Visualization

- Why visualize both Y and X?
  - Substantive: See which units you are talking about
  - Technical: adjust for skew
```{r}
debt%>%
  ggplot(aes(x=sat_avg))+
  geom_histogram()

debt%>%
  ggplot(aes(x=md_earn_wne_p6))+
  geom_histogram()
```
Step 3.) Multivariate Visualization

- Is this visualization good? Not really. We have some massive outliers.
```{r}
debt%>%
  ggplot(aes(x=sat_avg, y=md_earn_wne_p6))+
  geom_point()+
  geom_smooth(method="lm", se=F) # traditionally, this is how we'd plot a line of best fit. 
```

Step 4.) Regression

- Introducing the lm(formula, data) function
- Two inputs to care about:
  - formula: code for Y= alpha + beta*X
  - data: What is the data wer are using?
-formula is written as Y~X
  - R will calculate alpha and beta for us 
  - Just need to tell it what is the regressand and the regressor
  - The tilde is R's version fo the equal sign
  
Running the regression
```{r}
model_earn_sat <- lm(formula = md_earn_wne_p6 ~ sat_avg, data=debt)
summary(model_earn_sat)
```

**Another Example **
What is the relationship between admissions and future earnings?
  - There should be a negative relationship between admissions rate and future earnings
```{r}
glimpse(debt) # see variables

summary(debt%>%select(adm_rate, md_earn_wne_p6)) # NAs we have to remove. Affects how generalizable the data is.

debt%>% # univariate for adm_rate
  ggplot(aes(x=adm_rate)) +
  geom_histogram(bins=50)

debt%>% #univariate for med earnings
  ggplot(aes(x=md_earn_wne_p6))+
  geom_histogram()

#multivariate
debt%>%
  ggplot(aes(x=adm_rate, y=md_earn_wne_p6))+
  geom_point()+
  geom_smooth(method="lm", se=F)

# creating regression
model_adm_wage<-lm(formula = md_earn_wne_p6 ~ adm_rate, data=debt)
summary(model_adm_wage)
```
Conclusion: There is a negative relationship between admission rate and median earnings, indicated by the negative and statistically significant coefficient of admission rate.
