---
title: "2_27 Lecture"
author: "Alex Tomack"
date: "2023-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Gross ~ Score RMSE
```{r}
require(tidyverse)
mv<-read_rds("https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/5_Regression/data/mv.Rds?raw=true")
mv_analysis  <- mv %>%
  mutate(gross_log = log(gross),
         budget_log = log(budget))%>%
  select(gross_log, budget_log, score) %>%
  drop_na()

m<- lm(formula=gross_log ~ score, 
       data=mv_analysis)
summary(m)


# RMSE by CV

set.seed(123)
rmseRes<-NULL
for(i in 1:100){
  inds <- sample(x=1:nrow(mv_analysis),
                 size=round(nrow(mv_analysis)*.8),
                 replace=F)
  
  train<-mv_analysis%>%slice(inds)
  test<-mv_analysis%>%slice(-inds)
  
  m<-lm(formula=gross_log ~ score, data=train)
  
  e<-resid(m) # to get errors
  se <- e^2
  mse<-mean(se, na.rm=T)
  rmse<-sqrt(mse)
  
  rmseRes <- c(rmseRes, rmse)
}

mean(rmseRes)
```
Our RMSE is pretty much the same, implying that our model isn't susceptible to overfitting. 



# Multivariate regression

Yeeaaahhhh!
```{r}
m3<-lm(gross_log ~ gross_budget + score, data=mv_analysis)

summary(m3)
```
# RMSE Combined: Method 1

```{r}
set.seed(123)
rmseRes<- NULL
for(i in 1:100){
  inds<-sample(1:nrow(mv_analysis), 
               size=round(nrow(mv_analysis)/2),
               replace=F)
  train<- mv_analysis%>%slice(inds)
  test<- mv_analysis%>%slice(-inds)
  
  # NEW STUFF /////////////////////////////////
  
  mB<-lm(formula=gross_log ~ budget_log, train)
  mS<-lm(formula=gross_log ~ score, train)
  mC<-lm(formula=gross_log~ budget_log + score, train)
  
  rmseTmp<-test %>%
    mutate(pB = predict(mB, newdata=test), # budget
           pS = predict(mS, newdata=test), # score
           pC = predict(mC, newdata=test))%>% #combined
    summarise(rmseB = sqrt(mean((gross_log - pB)^2)),
              rmseS = sqrt(mean((gross_log - pS)^2)),
              rmseC = sqrt(mean((gross_log - pC)^2)))
  
  rmseRes<-rmseRes %>% bind_rows(rmseTmp)
}

rmseRes %>%
  summarise_all(mean) #takes a function and applies to every column in the data
```

# Visualize Result
```{r}
rmseRes%>%
  gather(model, rmse) %>%# instead of three columns, two columns. Wide to long data
  ggplot(aes(x=rmse, y= reorder(model, rmse)))+
  geom_violin()
```
# Categorical Predictor-- Bechdel Test 

- Do movies that pass the Bechdel Test make more money?
  - Theory: women are ~50% of the population. Movies that pass the test are more appealing to women. 
  - Hypothesis: Movies that pass the test make more money. 
  
When we run a regression, we find the a correlation that indicates movies that fail the test make more money. 
  
  - REMINDER correlation does not equal causation.
  - What might explain this pattern?
  - Budgets in a sexist Hollywood!
  - Movies that fail the test get larger budgets 
  - Budgets are positively associated with gross
- We want to control for the budget (analyzing two movies with the same budget) 
  - When we do this and run a multivariate regression, we see a negative coefficient on the bechdel_factorFail predictor for gross earnings. 