# Data Wrangling, 1/30

```{r}
rm(list=ls())
```

**What is data wrangling?**

- Where most of the work is done. Getting data ready for analysis
- Using excel is a mistake! It tries to make life easier by autocorrecting things. It will also convert things that look like dates INTO dates. 
  - REAL reason, it doesnt allow for replication
  
Three principles of avoiding data wrangling mistakes

- Replicability
- Understandability 
- Robustness

Big picture

- Load raw data into an RStudio script 
- Wrangle the data within the script

Why R?

- Checks all three principles 
- DON'T save workspace image
  - Always want a cleared workspace
  - Not clear workspace? type and run rm(list=ls())
    - Bisbee puts this rm(list=ls()) at the top of his code all the time. Just to protect in case you saved workspace. 
- R Scripts or RMD scripts should be:
  - Single purpose (but definitions can vary)
    - Like this file, it's organized around a single purpose (notes for this day)
  - Well-commented (make it easy for anyone to understand)
  - Iteratively built (constant bug checking)

New folder-- Output
```{r}
require(tidyverse)
MI_raw<-read_rds("https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/3_Data_Wrangling/data/MI2020_ExitPoll.rds?raw=true")

```
The process:

- Step 1: Look at the data
  - Methods:    
    - 1. Just look at the data. 
    - 2. Glimpse()
    - 3. Summary()
  - Types of Continuous variables
    - db1: numeric data stored with great precision. Can be manipulated with math
    - int: numeric data stored as integers. R typically treats db1 and int as interchangeable
  - Types of Categorical variables 
    - chr: string data, consisting of letters (and/or numbers). Cannot be manipulated with math 
    - fct: string data that is stored with a number. Typically used to define group membership.
  - Mixtures
    - lbl
  - Missingness
    - Two ways to indicate an observation is missing data 
      - 1. NA code
      - 2. Some bespoke code for "unit non-response" (often 9, or 99, or a large value ending in 9)
    ### looking for NA data USE SUMMARY
    ```{r}
    summary(MI_raw %>% select(LALVOTERID, SEX, AGE10, PARTYID, LGBT, QLT20))
    ```
    ### Looking for unit non-response USE COUNT
    ```{r}
    MI_raw %>%
      count(PARTYID) # have to knit to see the labels on each variable
    
    MI_raw %>%
      count(AGE10)
    ```
    - Reintroducing ggplot!
      - Only interested in single variable 
    - For now, need:
      - 1 aes
      - 2. geom_hist() or geom_bar()
    ```{r}
      MI_raw %>% 
      ggplot(aes(x=AGE10)) +
      geom_bar()
    ```
- Step 2: Wrangle the data
  - Why wrangle data?
```{r}
#what does 8.48 tell us? Half in category 8 and 9. BUT THIS IS INFLATED! We didn't remove the 99s in our data set. 
MI_raw %>%
  summarise(avg_age=mean(AGE10, na.rm=T)) 

#replace 99 with NA
MI_raw %>%
  mutate(AGE10_clean = ifelse(AGE10==99, 
                              NA, 
                              AGE10)) %>%
  select(AGE10, AGE10_clean)

# to assign the object so it actually exists, 
MI_raw <- MI_raw %>%
  mutate(AGE10_clean = ifelse(AGE10==99, 
                              NA, 
                              AGE10))
```
  - Converting to chr using Haven
    - There is a helpful package called haven which will extract these labels as factors 
    - install it but dont require it
  
```{r}
MI_raw %>%
  mutate(AGE10_clean_haven=haven::as_factor(AGE10)) %>%
  select(matches('AGE10'))

#then convert to character type
MI_raw %>%
  mutate(AGE10_chr = as.character(haven::as_factor(AGE10)))%>%
  select(matches('AGE10'))
```

## Plotting with labels

```{r}
MI_raw <- MI_raw %>%
  mutate(AGE10_chr= as.character(haven::as_factor(AGE10)))

MI_raw %>%
  ggplot(aes(x=AGE10_chr)) +
  geom_bar()
```
- Step 3: Analyze the data

Goals:

- 1. Understand variable types
  - Continuous, categorical
- 2. Identify missingness: either NA or unit non-response 
  - Unit non-response: observations who didn't provide information 
- 3. Identify skew or other phenomena that require wrangling

