# 1/23-- Intro to R, Part 3: Visualization

## Agenda

- Took Quiz 4. 

- First PSet will be posted today, fill in the functions according to the comments and respond to questions. **Due Midnight on Friday.**

## Lecture

**Opening data directly from internet**
```{r}
library(tidyverse)
## go to file location in GitHub, copy the link to the file

df<-read_rds('https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/2_Intro_to_R/data/sc_debt.Rds?raw=true') ##read_rds is different from readRDS. The latter is from base R, while the former is from tidyverse.
```

Remember the two camps of data science:

- Research and Prediction Camps
  - Research question: How might the admissions and SAT scores be related?
    - Theory: Admissions and SAT score should be negatively related. How do we test? **summarise() + filter()** 
    
```{r}
df %>% # Filters schools that have lower than 10% acceptance rates, then look inside the subgroup and find avg SAT 
  filter(adm_rate <0.1) %>%
  summarise(mean_sat=mean(sat_avg, na.rm=T)) 
```

```{r}
df %>% ## filters schools that are between 10 and 20% acceptance rate, then look inside subgroup and find avg SAT
  filter(adm_rate > 0.1 & adm_rate <0.2) %>%
  summarise(mean_sat=mean(sat_avg, na.rm=T))
```

Summmarise() + group_by()
```{r}
df %>% # arrange by the binary values of selectivity (0,1), then get their avg sat.
  group_by(selective) %>%
  summarise(mean_sat = mean(sat_avg, na.rm=T))
```


## Plotting Data

- Visualization in R uses ggplot
  - inputs aes(x, y, ...) 
- Theory: selective schools choose higher scoring students 
  - admissions rate goes on x, sat_avg goes on y
  
```{r}
df %>%
  ggplot(aes(x=adm_rate, y=sat_avg)) +
  geom_point() +
  geom_smooth(method="lm", se=F)
```

RQ: How might future earnings and SAT scores be related?

- Theory: SATs measure student ability

- Theory: Student ability is valued by the labor market

- Theory: firms pay more for students with higher SAT scores

- Hypothesis: earnings and SAT score should be generally positive

- Not be the case since SAT and earnings are caused by student ability or by socioeconomic position 

Variable assignment:

- x, predictor: avg_sat score

- y, predicted: earnings

```{r}
df %>% 
  ggplot(aes(x=sat_avg, y=md_earn_wne_p6)) +
  geom_point() +
  geom_smooth(method="lm", se=F)
```
```{r}
# What are the outliers?

df %>% # creates a new variable "outlier"
  mutate(outlier=ifelse(md_earn_wne_p6>100000,
                        instnm, # what is outputted given the condition 
                        NA)) %>%# What happens otherwise %>%
  ggplot(aes(x = sat_avg,y = md_earn_wne_p6,
             label = outlier)) +
  geom_point() + 
  geom_text()
```

## Categorical data

- What if we wanted to see which state has the most selective schools?

- Use group_by() and summarise()

```{r}
df %>% 
  group_by(stabbr) %>%
  summarise(adm_avg=mean(adm_rate,na.rm=T))%>%
  ggplot(aes(x=reorder(stabbr, adm_avg),
             y=adm_avg))+
  geom_bar(stat='identity') +#plot what I already gave you
  labs(title="Average Admissions by State",
       y="Admissions Rate",
       x= "State")

```
