# Beginning of Class

- Briefly worked, by ourselves, on selecting schools in Texas and finding the mean SAT.

- Remember to require(tidyverse) and RUN

  - Load data df<-("../data/sc_debt.Rcs")
    
    - "../" means go back one folder (up one level)
    
# Structured Data

- Standardized, well-defined, easily accessed 

# Unstructured

- Messy, organic. 

#Tabular Data Frame

- AKA tibble

- Rows: Units of observation 

  - People, places, things

- Columns: Variables of interest 

  - Beliefs, rainfall, likes, profits, etc. 
  
  
# Manipulating Data

Basic grammar of R is: object, %>%, verb

```{r}
object %>% #this is the object
  function() #this is the verb
```

Tidyverse has many verbs 

- filter(): will select rows based on some criteria

- select(): will choose certain columns

```{r}
df %>% 
  filter(instnm == "Vanderbilt University") %>% #Only select VU (case sensitive)
  select(instnm, adm_rate, sat_avg, md_earn_wne_p6) #select variables of interest
```

- arrange():

- filter()
```{r}
df %>%
  filter(adm_rate <0.1) %>% #Schools with AR <0.1
  arrange(desc(md_earn_wne_p6)) %>% #arrange from highest
  select(instnm, sat_avg, md_earn_wne_p6)
```
How good is Vanderbilt? Three data points

- Very  high SAT

- One of a few that select <10%

- However, we don't make quite as much as other T15s


