---
title: "2_6"
author: "Alex Tomack"
date: "2023-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
require(tidyverse)
mi_ep <- read_rds("https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/4_Uni_Multivariate/data/MI2020_ExitPoll_small.rds?raw=true")
```

Some light data science:

- The gender gap in Trump support
- Theory: Trump has expressed sexist views against women, therefore, women shoul dbe les slikely to support him
- Analysis: Compare support for Trump among men and women. 
  - Dependent variable: support for Trump
  - Independent variable: Sex
  - Multivariate: comparing support for Trump based on sex 
  
# Preparing Data Wrangling  

```{r}
Mi_final_small <- mi_ep %>%
  filter(preschoice=="Donald Trump, the Republican" | preschoice=="Joe Biden, the Democrat") %>%
  mutate(BidenVoter=ifelse(preschoice=="Joe Biden, the Democrat", 1, 0),
         TrumpVoter=ifelse(BidenVoter==1,0,1),
         AGE10=ifelse(AGE10==99,NA, AGE10))
```

# Analysis 
```{r}
Mi_final_small %>%
  count(SEX, preschoice) %>% # shows pres choice for each gender
  mutate(prop_overall=n/sum(n)) #shows prop of total voters, but we want prop of total women 
```

# Analysis of JUST women| One Way Conditional Mean 
```{r}
Mi_final_small %>%
  count(SEX, preschoice) %>% # shows pres choice for each gender
  mutate(prop_overall=n/sum(n)) %>%#shows prop of total voters
  group_by(SEX) %>% 
  mutate(tot_resp = sum(n)) %>% # sums each gender
  mutate(prop_gender=n/tot_resp) %>% # now, proportion by gender
  select(SEX, preschoice, prop_gender) # make it a bit more readable
```

So have we answered our research question? Is there a gender gap in Trump voters?

- Looks like it, but is it really conclusive? 

# Additional Theorizing 

- The strength of the theorized relationship (the gender gap) may vary be age
  - Younger women might be more offended by Trump's casual sexism
  - Older women might be more inured to Trump's casual sexism
- Theory: the "gender gap" will be larger among younger voters
  - But also recognize that younger Americans are generally more progressive, meaning that both younger men and women are more offended by Trump's casual sexism.

# Two-Way Conditional Mean
```{r}
toplot <- Mi_final_small %>%
  group_by(SEX, AGE10) %>%
  summarise(prop_trump=mean(TrumpVoter, na.rm=T),.groups="drop") %>%
  mutate(prop_trump=round(prop_trump, digits=2)) %>%
  spread(key=SEX,value=prop_trump) %>% # spreads the Sex value across multiple columns.
  rename(Male=`1`, Female=`2`) %>%
  mutate(gender_gap=Female-Male)
```

# plotting data toplot
```{r}
toplot %>%
  ggplot(aes(x=AGE10)) +
  geom_bar(aes(y=Male), fill="forestgreen", stat="identity", alpha=.5) +
  geom_bar(aes(y=Female), fill="skyblue", stat="identity", alpha=.5)
# but this is very confusing, so lets just plot the gender gap we created
```

```{r}
toplot %>%
  ggplot(aes(x=AGE10, y=gender_gap)) +
  geom_bar(stat="identity")
```

# Spread command

- Check gif in the slides to see exactly how it works, is SUPER useful 

# Conditional Categorical Analysis

Looking at reasons for voting for either candidate, we can't answer any questions but we rather generate them

Can also do it by sex

# Theorizing 

- Research Question: Why do polls under-predict Trump more than Biden 
  - Weary Trump voters-- if they're approached by CNN, NYT, etc., Trump voters likely wont respond. 
  - Trump voters might be embarrassed to be Trump voter. 
  - Fake news lol 
    - "Any negative polls are fake news, just like the CNN, ABC, NBC polls in the election. Sorry, people want border security and extreme vetting."

