---
title: "Intro to R"
subtitle: "Lecture 1-- Installing R"
author: "Alex"
institute: "Vanderbilt University"
date: "2023-01-11"
output: html_document
---
```{r}
require(tidyverse)
```
# Lecture 1

## Organization

"knit" compiles code-- outputs in a nice format

"#" Heading

"##" subheading

"###" subsubheading 

- Have to include spaces after tags for it to register 

1. Numbered list 

- bulleted list


## Formatting

**bold font**

*italic*

`code font` with backtick (tilde key)

## How to actually write code

``` {r}
2+2
```

## How R works 

- Object Oriented Languge (OOL)

- Objects are created with the <- command

**Assignment operator**:
``` {r}
object <- 2+2

object
```

## Different types of objects

- An element is a single value (number, letter, word)

- A vector is a list of values 

  - We create a list with the `c(value1, value2, value3)`
  
```{r}
a_vector <- c(1,2,3)
a_vector
```

- A list: stores multiple different value/elements/vectors

```{r}
a_list <- list(this_is_4 = 2+2,
               this_is_text = "hello",
               this_is_a_vector = runif(n = 3, min = 0, max = 10)
               )
a_list
```


## This section is about comments

```{r}
#This line creates a vector
a_vector <- c(10,29,100)
a_vector
```

## How RStudio Works 
- RStudio is deeply interactive-- we can run individual code chunks
  - We can also highlight specific lines of code by highlighting it and pressing ctrl+enter
  
  
## Give it a try
``` {r}
# a is mulitplying 3 and 5
a <- 3*5
# b s a vector of the following values
b <- c(10,21,43,87,175)
# c is a vector output of the following arithmetic
c <- a-b 
c
```
## Functions and Packages 
- What are packages? 
  - Functions that someone else wrotes
- R has many functions already installed
  - These are known as "Base R", for example, sum()
```{r}
sum(object1)
```
- Installing packages
  - To install tidyverse, for example, "install.packages("tidyverse")
- Loading packages in R, type:
```{r}
require(tidyverse)
#or
library(tidyverse)
```