---
title: "2_8 Lecture"
author: "Alex Tomack"
date: "2023-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)
poll<-read_rds("https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/4_Uni_Multivariate/data/Pres2020_PV.Rds?raw=true")
  
poll %>%
  select(Trump,Biden)

poll<-poll%>%
  mutate(Trump=Trump/100,
         Biden=Biden/100,
         margin=round(Biden-Trump,2)) 
```
Why do Polls under-predict Trump more than Biden?

- Unrepresentative sample (how were respondents contacted?)
- Small samples
- Shy Trump voters 
- Timing (closer to election)

Theory #1: Does the "mode" of the survey  matter?

- I.e., if you only call people on landlines, who do you reach?
- And how might they differ form the general population
- Assumption #1: Younger people do not use landlines, meaning that surveys which rely on random digit dialing (RDD) will get disproportionately older respondents.
- Assumption #2: younger voters are more progressive, making them less likely to support Trump.
- Theory: Surveys that use RDD will find more support for Trump than Biden

How do we analyze this?

- plot the Biden-Trump vote margin by mode type

Soft Rules of Visualization 

- Variable type informs visualization 

1.) Univariate
  - Categorical data: geom_bar()
  - Continuous data: geom_histogram() or geom_density()

2.) Bivariate
  - Categorical x Categorical: geom_bar()
  - Binary x Continuous: geom_histogram() or geom_density()
  - Categorical x continuous: geom_boxplot() or geom_violin()
  - Two continuous: geom_point()
  
- For now, just focus on IRV/Online versus Live Phone - RDD
- Since margin is a continuous variable, use geom_histogram

# Bivaiate vis: binary x continuous
```{r}
poll %>%
  filter(Mode =="IVR/Online" | Mode=="Live phone - RDD") %>%
  ggplot(aes(x=margin, color=Mode))+
  geom_density()
```
# Bivariate vis: categorical x continuous
```{r}
poll %>%
  filter(Mode !="IVR" & Mode!="Online/Text" & Mode != "Phone - unknown") %>%
  ggplot(aes(x=margin))+
  geom_boxplot()
```
#bivariate vis: cont x cont 3rd method besides alpha and jitter 
```{r}
poll%>%
  ggplot(aes(x=Biden, y=Trump)) +
 # geom_point(alpha=0.3)+ # approach 1 to revealing multiple data at same point.  alpha shows how many polls at the same point. 
 # geom_jitter()+ # approach 2, shifts the points a tiny bit so we can see each poll individually
  geom_point()+
  geom_smooth(method="lm")
```
#bivariate vis: cont x cont 
```{r}
poll%>%
  count(Biden, Trump) %>%
  ggplot(aes(x=Biden, y=Trump, size=n)) +
  geom_point()+
  geom_smooth(method="lm")
```
Theory 
 
- These reslts indicate that polls whcih rpedict greater support for Biden also predict greater support for Trump
  - Is this consistent with the theory?
  - Recall that Biden-biased polls should underpredict Trump support and overpredict Biden support
  - in the data, this would suggest a negative relationship
  - but we find a positive relationship
- Inconsistent with the theory, but raises another question 
- Why do polls that underpredict support for Biden also underpredct support for Trump?
  - Third party bias? Polls bias against 3rd party candidates
  - Timing of poll? Fewer uncertain responses closer to election.
  
To test the timing of polls theory, let's look at how predicted support varies over time 
```{r}
poll %>%
  select(StartDate, EndDate, DaysinField) 
```

Working with dates-- as.Date() dunction converts characters to 'date' class variables
```{r}
d1<-as.Date("02/08/2023", format="%m/%d/%Y")
midterm <-as.Date("03/08/2023", format="%m/%d/%Y")
midterm-d1

poll<-poll %>% 
  mutate(StartDate=as.Date(StartDate, format="%m/%d/%Y"),
         EndDate = as.Date(EndDate, format="%m/%d/%Y"))

poll %>%
  ggplot(aes(x=StartDate, y=margin))+
  geom_point() +
  geom_smooth(method="lm")
```


Look at Each point individually

```{r}
poll %>%
  ggplot() +
  geom_point(aes(x=StartDate, y=Biden),color='Blue') +
  geom_point(aes(x=StartDate, y=Trump), color='Red')
```
