---
title: "Univariate Analysis"
subtitle: "Lecture Notes 2/1/23"
author: "Tomack"
institute: "Vanderbilt University"
date: "2/1/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE,warning = FALSE,message = FALSE)
```
Scope

- How to analyze a single variable
- Scientifically, theories contain more than one variable?
  - Relationship between education and wage; gender and voting
- How can we put together a theory with a single variable i.e. no context??

Still, uni analysis is essential

- Informs how we wrangle data (practical)
- Allows us to generate hypotheses (scientific)

```{r}
#Load data on players in the 2018-2019 season. 
require(tidyverse)
nba <- read_rds("https://github.com/jbisbee1/DS1000_S2023/blob/main/Lectures/4_Uni_Multivariate/data/nba_players_2018.Rds?raw=true")
```


Thinking like a scientist 

- What questions do we have? What hypotheses might we want answered?
- Take a look at the pts variable 
  - what does this measure? What is its variable type?
```{r}
glimpse(nba%>% select(pts)) # stored as a double (dbl), so continuous variable. 
```

But how do we actually analyze a single variable?

- summarise the mean, median

```{r}
nba %>%
  summarise(mean_pts=mean(pts, na.rm=T),
            med_pts=median(pts, na.rm=T)) #mean>med, right-skewed
```

OR

- summary(nba$pts)

```{r}
summary(nba$pts) # $ JUST selects that one column 
```
In English:

- There is at laest one player who didnt score at all 
- At least one player scored 2818 points
- 25% of players scored less than 115 points
- What do decimals mean?
  - you can't score half a point in bbal.. it's just the nature of the math. 
  
Visualization 

- We could try to remember the statements, OR we could graph them!

```{r}
# Univariate visualization using geom_hist(). ONLY variable will be on the x axis

nba %>%
  ggplot(aes(x=pts))+
  geom_histogram(bins=150) +#default bins = 30, can change using bins=...
  geom_vline(xintercept=quantile(nba$pts,c(.25, .5, .75))) # draw a vertical line that intercepts the x-axis using the quantile function-- look in the nba data set and take the pts variable and plot the data from pts AT the first quartile, median, and 3rd quartile
```
```{r}
quantile(x=nba$pts, .27) # takes the value of the variable at the percenctile given
#or
quantile(x=nba$pts,c(0.25, 0.5, 0.75)) #take the values of the variable at the vector of percentiles given

nba%>%
  ggplot(aes(x=pts))+
  geom_histogram()+
  geom_vline(xintercept=750) # just showing that the 750 in xintercept=750 can be interchanged with a function like quantile(x=nba$pts,c(0.25, 0.5, 0.75)) 
```
These plots are BAD though! If someone were to walk in and look at the plot, could they tell what it means?

- Need to assign labels and titles. What does pts mean? What is count *counting*??

Labs() function allows us to do this
```{r}
nba %>%
  ggplot(aes(x=pts)) +
  geom_histogram() +
  geom_vline(xintercept=quantile(nba$pts, c(.25,.5,.75)))+
  labs(title = "Total Points Scored",
       subtitle="NBA Players 2018-2019 Season",
       x = "Total Points Scored",
       y = "Number of players") #adding labels
```

Visualization informs science

- Looking at the data can help generate research questions, theories, and hypotheses
  - Question: Why do some players not score any points?
  - Theory: Players need minutes to score points
  - Hypothesis: The number of points a player scores should be positively correlated with their minutes
  
```{r}
summary(nba$minutes)
```

Lets plot minutes 

```{r}
nba %>%
  ggplot(aes(x=minutes)) +
  geom_histogram() +
  geom_vline(xintercept=quantile(nba$minutes, c(.25, .5, .75)))+
  labs(title=" Minutes by PLayers",
       subtitle="2018-2019 NBA Season",
       x = "minutes played",
       y = "Number of Players")

```

Other Variables

- Thus far, pts and minutes are both doubles (dbl)
- What about other variable types like categorical?
  - Categorical variables are:
    - Mutually exclusive: Observations can only be in one category 
    - Exhaustive: Every observation is assigned to a category

Categorical Variables

- Categorical variables can be divided into the followign sub-types 
- **Ordered**: There is a sensible order (i.e. education)
  - Should be arranged intuitively (i.e. LTHS, HS degree, Some college)
  - To summarize: calculate the proportions for each category
  - If there are too many categories, use the mode 
  - To visualize, **geom_bar()**
  - **Ordered-Binary** is a sub type
    - Arranged in intuitive order (is not a rookie/is rookie)
- **Unordered**: No sensible order of categories (i.e. major degree)
  - Order by most commonly occurring categories
  - **Unordered Binary**: No sensible order and only two levels (i.e. edible)
    - To summarize, convert to a [0,1] and find mean 
    - Visualize with geom_bar()
- Meaningfully different from continuous variables
- In general, continuous uses geom_histogram or geom_density, categorical uses geom_bar
- fct is a class that is unique to R
  - meant for ordered categorical variables
  - fct stores the order and assigns a numeric value + a definition 
  - Most of the time, better to store as a chr (but not always)
  
Variables
- R may store categorical variables as chr, fct, lgl, or even dbl 
- continuous variables typically stored as int or dbl
- Up to the data scientist to determine 
  - simple process: 
    1. look at a few observations and make a guess about the variable type
    2. create a plot or table based on that guess
    3. if the result is sensible, proceed. Otherwise, go back to #1
    
    
In Practice

- Let's look at field goals (fgm)
- What type of variable should this be?
  - Technically not continuous, since it can't be divided into fractions, but we don't care.
  - We just want to make sure it is not a categorical variable 
The Process:
- Step 1: Look
```{r}
summary(nba$fgm) # we looked, so lets try plotting

nba %>%
  ggplot(aes(x=fgm))+
  geom_histogram(bins=100) #geom histogram since continuous, BUT can also do a DENSITY!

nba%>%
  ggplot(aes(x=fgm))+
  geom_density() #draws the same shape, just depends on the type of visualization
```
Step 3: Evaluate
- blah blah blah whatever we conclude 

Other variables: Use the process

```{r}
nba%>%
  select(slugTeam) #clearly a categorical variable

nba %>%
  ggplot(aes(x=slugTeam))+
  geom_bar()
```

  